Using Stochastic Modeling to Predict Long Timescale Transport Behavior of Solutes in an H\textsubscript{II} Phase Lyotropic Liquid Crystal Membrane

### 2019-08-27 Ben_Manuscripts/stochastic_transport/Outline.tex
Outline updates

  Unfortunately, the timescales that we can simulate with MD are insufficient to be
  able to make well-converged predictions of macroscopic transport properties 
  traditionally used to characterize membranes in the lab. \\
  \begin{itemize}
    \item However, if we use descriptive stochastic models that can capture solute
    dynamics, then we could project long timescale behavior in addition to gaining
    a deeper understanding of solute behavior on short timescales.
  \end{itemize}

  \noindent Researchers have developed a rigorous theoretical foundation for describing
  the motion of particles exhibiting non-Brownian, or anomalous, transport behavior.
  \begin{itemize}
    \item The tools introduced by fractional calculus are instrumental to this theory. 
    \item They allow us to generalize the normally linear diffusion equation using fractional
    derivative orders, providing descriptions of a much more diverse set of behavior. % diverging from the relatively simple Brownian model.
%    \item They facilitate proof of the relationship between the observed growth of a
%    particle's mean squared displacement and underlying dwell times 
%    \item In the absence of an external potential or complicated boundary conditions, 
%    one can define random walk based on independent hop length and dwell time 
%    distributions and arrive at any type of diffusive behavior. 
  \end{itemize}

  % BJC1: moved from methods
  \noindent Three well-known classes of behavior leading to anomalous subdiffusion are 
  continuous time random walks (CTRW), fractional Brownian motion
  (FBM) and random walks on fractals (RWF).\cite{meroz_toolbox_2015}
  \begin{itemize}
    \item A CTRW is characterized by a distribution of hop lengths and 
    dwell times, where each each hop is characterized by independent random draws from 
    each distribution.\cite{montroll_random_1965,morrin_three_2018}
    \item FBM is common in crowded, viscoelastic environments where each jump comes 
    from a Gaussian distribution but is anti-correlated to its previous 
    step.~\cite{mandelbrot_fractional_1968,jeon_fractional_2010,banks_anomalous_2005}
    \item An RWF is imposed by a system's geometry. Systems with tortuous pathways and dead
    ends cause anti-correlated motion.\cite{meroz_toolbox_2015,neusius_subdiffusion_2008}
    \item The processes described above can happen alone or in combination. 
    \item Meroz and Sokolov suggested a decision tree for determining which subdiffusive
    mechanisms apply to a system, adapted in Figure X. 
    \item Of course, more complicated variations of these models can arise which we 
    will incorporate as needed.	  % BJC1: e.g. fractional levy motion or langevin with levy noise
  \end{itemize}

@@

  While elegant, pure anomalous diffusion theory may in some cases give only a qualitative
  understanding of transport mechanisms.
  \begin{itemize}
    \item Complex temporal relationships in between hops may have a significant effect on MSD
    \item Inhomogeneity of a medium or multiple types of interactions may lead to a number of
    distinct diffusive regimes. 
    \item In this case, a state based model may be useful. 
  \end{itemize}

  We can use the observations of transport from our previous work in order to create 
  state-based models.
  \begin{itemize}
  	\item We know that dynamics differ between solutes moving inside the pores and in the tails.
  	\item We can fit the above models to our data as a function of location.
  	\item Alternatively, we can abandon the anomalous diffusion theory and adapt the 
  	methods of Markov state models.
  \end{itemize}

  Markov state models blah blah blah.
  \begin{itemize}
    \item Markov state models are popular with researchers of proteins which have 
    many dynamical states occurring over multiple time scales.
  \end{itemize}

  We will extend the Markov state models framework to include dynamics between jumps.

@@

  We can use the observed mechanisms in order to choose a stochastic model of
  sufficient complexity One can project long timescale solute behavior by modeling transport as a
  stochastic process.
  %MRS: as above, simple diffusion is stochastic, just not very complex.  So we want to add appropriate complexity.
  \begin{itemize}
    %MRS: best talk about these mechanisms in terms of the
    %mathematical models they capture.  But I don't know that it's
    %necessary to review all of them.

    %It may be a better approach to start from the features of the
    %problem (i.e. certain mechanisms) and choose the model (of the
    %existing choices) that is appropriate for those mechanisms.
    %I.e. diagnose first, then choose, rather than just applying a
    %list of models useful in other field's timeseries. Let me know if that makes sense;
    %I think how the methods applied are selected is a key to solve the problem.

    \item Various disciplines study this type of problem in a number of ways.
    \item In the simulation literature, most start by measuring mean squared 
    displacement (MSD), as we did previously.  % This doesn't quite fit the thesis
    \item Due to the short time scales most people can feasibly simulate, the MSD 
    does not give a complete picture of what happens long-term, other than hinting
    at types of mechanisms that could lead to the curve's shape.
    %MRS: I'd say that MSD is too simple compared to what is going on mechanistically, but it's fundamentally the same idea.  
    \item mean first passage time?
    \item Markov state models are popular with researchers of proteins which have 
    many dynamical states occurring over multiple time scales.
    \item The anomalous diffusion literature explains the shape of MSD curves based
    on rigorous mathematical representations of transport mechanisms. This field is
    heavily used in analysis of single particle tracking experiments.
    \item Finally, the financial literature tackles similar problems from a more 
    general time series perspective. Their primary focus is towards predictions of
    stock market returns and volatility but their methods are widely applicable to
    modeling time series that exhibit a range of behavior.
  \end{itemize}

  In this work, we explore two approaches to model the long timescale behavior
  of the four solutes fasted moving solutes studied in our previous work.
  \begin{itemize}
    \item Specifically, we study methanol, urea, ethylene glycol and acetic acid
  	\item Our first approach uses Markov state models (MSMs) 
  	\item The second approach is based on the anomalous diffusion literature.
%  	\item The third uses an infinite hidden Markov model (iHMM), an unsupervised machine
%  	learning algorithm which groups solute behavior into different dynamical states.
  	% BJC: somewhere emphasize that this last approach is powerful since it doesn't 
  	% require anything different in systems with more complex geometries (i.e. QI)
  \end{itemize}

@@

  We studied transport of solutes in the H\textsubscript{II} phase using an
  atomistic molecular model of four pores in a monoclinic unit cell with 
  10 \% water by weight. 
  \begin{itemize}
    \item Approximately one third of the water molecules occupy the tail region 
    with the rest near the pore center.
    \item We chose to study the 10 wt \% water system because solutes move 
    significantly faster than in the 5 wt \% system studied previously.
    \item Appropriate stochastic modeling requires that solutes sample the 
    accessible mechanisms with representative probability.  %BJC1: incorporated suggested phrasing
  \end{itemize}

@@

  \subsection{Markov State Models}\label{method:MSMs}  % BJC: this section first in hopes that sfbm works better

  A Markov state model (MSM) decomposes a time series into a set of discrete states
  with transitions between states defined by a transition probability matrix, T,
  which describes the conditional probability of jumping to a specific 
  state given the previously observed state.~\cite{pande_everything_2010,wehmeyer_introduction_2018}
  \begin{itemize}
    \item In the context of molecular simulations, MSMs are frequently used
    to study systems with slow dynamics, such as protein folding.~\cite{snow_how_2005,chodera_automatic_2007} % might move to intro
    \item Researchers typically aim to come up with a low dimensional 
    representation of the system based on features which preserve the process
    kinetics. This facilitates the identification of discrete states from
    which T is generated.
    \item Software packages such as MSMbuilder~\cite{beauchamp_msmbuilder2:_2011}
    and pyEMMA~\cite{scherer_pyemma_2015} provide work flows capable 
    of featurization and dimensional reduction.
  \end{itemize}

  %MRS: probably write it as a 2x4 table.
  In this work, we define a total of 8 discrete states based on the 3 trapping
  mechanisms observed in our previous work. 
  \begin{itemize}
  	\item Therefore, there is no need to subject our trajectories to any reduction
  	in dimensionality. 
  	\item However, we will adopt the validation techniques used in the packages
  	mentioned above.  
  	%BJC: not sure the best way to lay these out
  	\item The states we've chosen include all combinations of trapping mechanisms 
  	in the pore and out of the pore:
  	\begin{enumerate}
  	  \item Trapped in tails
  	  \item Trapped in tails and hydrogen bonding
  	  \item Trapped in tails and associated with sodium
  	  \item Trapped in tails, hydrogen bonding and associated
  	  \item In pores
  	  \item In pores and hydrogen bonding
  	  \item In pores and associated with sodium
  	  \item In pores, hydrogen bonding and associated
  	\end{enumerate}
  	\item These choices assume that there are no significant kinetic effects resulting
  	from solute conformational changes or pore size fluctuations, an assumption 
  	that may be relaxed in future work.
  \end{itemize}

  % BJC: does it make sense to use Dirichlet Process as a conjugate prior for
  % getting error on transition matrix?

  \noindent We constructed a state transition probability matrix based on observed solute
  trajectories.
  \begin{itemize}
    \item Using methods described in our previous work, we determined 	
    which, if any, trapping mechanisms affected the solute at each time step, and
    assigned the observation to a specific state according to the definitions 
    above.~\cite{coscia_chemically_2019}
    \item Based on the current and previous state observation, we incremented the
    appropriate entry of an $n x n$ count matrix by 1, where $n$ is the number of states.
    \item For example, if we observed state 1 followed by a transition to state 3,
    we increment the $(1, 3)$ entry of the count matrix by 1.
    \item We generate the transition probability matrix from the count matrix by 
    normalizing the entries in each row so that they summed to unity.
  \end{itemize}

  \noindent We recorded the $z$-direction displacement at each time step in order to construct
  individual emission distributions for each state and transition between states.
  \begin{itemize}
    \item While the dynamics of the states themselves are important, the 
    majority of observations involve transitions between states, so properly modeling
    the transition dynamics is even more important.
    \item This results in 64 distinct emission distributions. Some are far more
    populated than others. % BJC: not sure if the emission distributions are reversible. i.e. if 1 --> 2 transition has same emission distribution as 2 --> 1. It might be required to be Markovian since transition matrix is symmetric. Not sure.
    %\item The shape of the distributions are symmetric and heavy-tailed. 
  \end{itemize}

  % MRS: this below sort of goes beyond standard use of MSM's though, right? where everything is 
  % usually constant-rate expoential processes? Should be explicitly mentioned.N
  %BJC1: yes
  \noindent We modeled the emission distributions as L\'evy stable distributions.  % cauchy is a case of a Levy stable distribution. The tails are too heavy though.
  \begin{itemize}
    \item For independent and identically distributed (iid) random variables, the 
    generalized central limit theorem guarantees convergence of the associated 
    probability distribution function (PDF) to a L\'evy stable PDF. \cite{klages_anomalous_2008}
    % BJC: ``For sums of independent, iid random variables with proper normalization to the
%    sample size, the generalized central limit theorem guarantees the convergence of
%    the associated PDF to a L\'evy stable PDF even though the variance of these random
%    variables diverges." \cite{klages_anomalous_2008}  % BJC: reword
    \item The characteristic equation describing the Fourier transform of a L\'evy 
    stable PDF is given below:
    \begin{equation}
    p_{\alpha, \beta}(k;\mu,\sigma) =\exp\left[i\mu k - \sigma^{\alpha}|k|^{\alpha}\left(1 - i\beta\frac{k}{|k|}\omega(k, \alpha)\right)\right]
    \end{equation}
    \item Where $\omega$ is defined with something that is complicated to format and can be 
    added later if I keep these formulas in. % https://tex.stackexchange.com/questions/32140/how-to-write-a-function-piecewise-with-bracket-outside
    \item $\alpha$ is the index of stability or L\'evy index, $\beta$ is the skewness 
    parameter, $\mu$ is the shift parameter and $\sigma$ is a scale parameter.
    \item The most familiar case, and 1 of 3 that can be expressed in terms of elementary
    functions, is the Gaussian PDF ($\alpha$ = 2).
    \item The more general family of L\'evy stable distributions allows
    greater flexibility in defining the observed emission distribution PDFs.
    \item In the case that the empirical emission distributions are iid, sequential draws
    from a L\'evy stable PDF defines a L\'evy process. 
    \item We can relax the iid assumption and maintain the flexibility of L\'evy stable 
    probability distributions if we instead consider the dynamics to be governed by a
    non-Gaussian Ornstein-Uhlenbeck (OU) process.\cite{barndorffnielsen_non-gaussian_2001} % BJC: replaces Wiener process with Levy process
    \item OU processes contain temporal dependence but the marginal observation distributions
    are consistent with the underlying stochastic driver.~\cite{taufer_simulation_2009}
    \item We will discuss the implications of each.  
    %MRS: does the process described above include the motion within the state? That is something I feel that the
    %standard MSM doesn't include either, since they are typically interested just in the distribution of states in equilibrium, not the diffusion. 
  \end{itemize}

  L\'evy stable distributions often have heavy tails and an undefined variance 
  which can give rise to arbitrarily long particle displacements.
  \begin{itemize}
  	\item We do not observe hops longer than x nm in our simulations which 
  	emphasizes that these distributions are only an approximation.
    \item We avoid enormous and unrealistic hops problem by truncating the tails
    of the distribution. %BJC: can add some citations about truncated Levy distributions which are used frequently in modeling
    \item Since solutes in our system cannot hop further than the length of our 
    simulation unit cell, we truncate the tails of the distribution at this
    magnitude.
    % BJC: Is this actually the right way to do this? I am looking at unwrapped coordinates,
    % so theoretically a particle can move as far as it wants between frames.
    % BJC: The cut-off choice is pretty important and has a significant effect on MSD. 
    % For example, if I truncate at the unit cell length I reproducably get a final MSD of ~ 66 nm^2 (1000 ns stochastic simulation)
    % If I truncate the tails at double the unit cell length, I get a final MSD of ~107 nm^2
    % The uncertainty also increases (since variance diverges for these distributions)
    % So maybe it makes the most sense to truncate based on largest observed hops.
  \end{itemize}

  We simulated realizations of the stochastic process using the probability transition
  matrix and emission distributions.
  \begin{itemize}
    \item For each trajectory simulated, we chose an initial state randomly from a 
    uniform distribution. %BJC: could construct initial distribution but I'm not sure how reliable that is based on 24 solutes that are started in the pore center. (need to chop off beginning 5 ns or so to get rid of that initial dependence)
    \item We randomly drew subsequent state transitions and corresponding emissions 
    from the rows of the probability transition matrix and the appropriate emission
    distribution respectively.
  \end{itemize}

  %MRS: you intend this to be a separate model than the above?  If so, make clearer.
  %BJC1: renamed, is that what you meant?
  \subsection{The Anomalous Diffusion Model}\label{method:model_sFBM}

  Solutes in this system exhibit subdiffusive behavior, a type of anomalous diffusion.
  \begin{itemize}
  %MRS: you would probably want to be clear what the mechanistic reasons for for anomalous diffusion: 
  %want to talk about why, given the mechanisms you would consider. 
  %BJC1: could go through flow chart
  	\item During an anomalous diffusion process, the mean squared displacement (MSD)
  	does not grow linearly with time, rather it is of the form:
	\begin{equation} 
	\langle x^2(t) \rangle = K_{\alpha}t^\alpha
	\label{eqn:msd_form}
	\end{equation} 
	where $\alpha$ is the anomalous exponent and $K_\alpha$ is the
	generalized diffusion coefficient.
	\item A value of $\alpha < 1$ indicates a subdiffusive process, while values of
	$\alpha = 1$ and $\alpha > 0$ are characteristic of Brownian and superdiffusive
	motion respectively.
  \end{itemize}

@@

  % BJC: Should this be in the results? Lots of figures to show in support of the claims below
  %MRS: can you provide mechanistic reasoning for the below? Then can show the claims for the reasonable possibilities.
  \noindent We believe that solutes in the system studied here exhibit subordinated 
  fractional Brownian motion (sFBM) where the parent process is FBM and the 
  leading process is a CTRW. 
  \begin{itemize}
  	\item The ensemble-averaged MSD differs from the time-averaged MSD, which
  	is indicative of non-ergodicity, a trait inherent to CTRWs but not FBM or RWFs.~\cite{thiel_weak_2014}  % BJC: need a figure to show this
  	\item We also observe non-stationary $z$-coordinate traces of each solute's
  	center of mass (COM). %BJC: figure of a z-coordinate trace in main text or in supporting info
  	\item For a pure CTRW, the time-averaged MSD should be linear.
  	~\cite{neusius_subdiffusion_2008,meroz_subdiffusion_2010}
  	\item However, a typical time-averaged solute MSD is sublinear (see supporting
  	information), which suggests that there is another underlying subdiffusive mechanism.
  	\item The hop lengths recorded after each dwell period are anti-correlated (See supporting information)
  	\item Given the viscoelastic nature of the monomers in our system, we believe
  	the hop lengths can be modeled with FBM. 
 	\item For subordinated FBM, it can be shown that
  	\begin{equation}
  	\langle x^2(t) \rangle \simeq t^{\alpha\beta}
  	\end{equation}
  	where $\alpha$ is the anomalous exponent characteristic of the leading CTRW process
  	and $\beta$ is the anomalous exponent characteristic of the parent FBM process. 
  \end{itemize}

  \noindent We can characterize a CTRW process using the parameters which describe its
  dwell time and hop length distribution.  % This is different from the emissions of the markov state model where every single step is used to construct the distribution. 
  \begin{itemize}
	\item We used the \texttt{ruptures} python package in order to identify
	changepoints in solute trajectories.\cite{truong_ruptures:_2018} (See Supporting
	Information for more details on chosen parameters. i.e. type of cost function, 
	cost function penalty tolerance, number of dimensions used)
	\item We used the corresponding hop lengths and dwell times between break points
	to construct empirical distributions.
  \end{itemize}

  For solutes in our system, the distribution of hop lengths appears to be
  well-represented by a Gaussian distribution.~\cite{metzler_random_2000,
  metzler_anomalous_2014,neusius_subdiffusion_2009}  % Note that this is different from the distribution of all increments (which is heavy-tailed)
  \begin{itemize}
	\item We are most interested in the standard deviation, $\sigma$, of the 
	hop length distribution.
  \end{itemize}

@@

  % BJC: Is it worth showing performance of both types of models? one mode versus two mode?
  % Probably worth showing that 1 state fails
  % BJC: also not sure if there is a better word than mode. I stayed away from 'state' since
  % it's used heavily in the MSM section.
  % BJC: I haven't created the 2 mode model yet, but that'll only take a few hours. I'll
  % do that after I get BCC equilibrations running
  In general, we observe different dynamical behavior when solutes move inside the
  pore versus in the tail region.

  %MRS: could present the two models sequentiually.  Talk about one COULD have multiple modes, then in results present 1, then 2.

  \begin{itemize}
    \item This inspired two models of varying complexity.
    \item We created a simple, single mode model with a single $\alpha$, $\sigma$ and $H$
    parameter fit to each solute.
    \item Our second, two mode model assigns a set of parameters to each of 2 modes based
    on the solute's radial location.
    \item Solutes in mode 1 are in the pore region defined as less than 0.75 nm from any
    pore center and all else are in mode 2, the tail region. 
    \item We determined this cut-off and described how we calculated radial 
    distance from the pore center in our previous work~\cite{coscia_chemically_2019}
  \end{itemize}

@@

  % BJC: can ignore this section for now. 
  \subsection{The Infinite State Hidden Markov Model}\label{method:iHMM}  %BJC: potentially a separate paper

  Hidden Markov models (HMMs) are a useful and widely used technique
  for modeling sequences of observations where the probability of the next observation
  in a sequence depends only on a previous unobserved, or hidden, state.~\cite{beal_infinite_2002}
  \begin{itemize}
    \item In the context of our simulations, the observations correspond to 
    the center of mass coordinates of the solutes versus time, and the states
    correspond to the dynamical behavior which give rise to those types
    of observations.
    \item Unfortunately, standard HMMs require the number of hidden states to be known
    a priori.
    \item One can partially overcome this by testing a range of numbers of 
    hidden states and determining which is the best representation of their
    data.
  \end{itemize}

>>>>

  Unfortunately, the timescales that we can simulate with MD are insufficient to be
  able to make well-converged predictions of macroscopic transport properties 
  traditionally used to characterize membranes in the lab. \\
  \begin{itemize}
    \item However, if we use descriptive stochastic models that sufficiently capture solute
    dynamics, then we could project long timescale behavior in addition to gaining
    a deeper understanding of solute behavior on short timescales.
  \end{itemize}

  \noindent Researchers have developed a rigorous theoretical foundation for describing
  the motion of particles exhibiting non-Brownian, or anomalous, transport behavior.
  \begin{itemize}
    \item The tools introduced by fractional calculus are instrumental to this theory. 
    \item They allow us to generalize the normally linear diffusion equation to fractional
    derivative orders, providing descriptions of a much more diverse set of behavior. \cite{klages_anomalous_2008} % diverging from the relatively simple Brownian model.
%    \item They facilitate proof of the relationship between the observed growth of a
%    particle's mean squared displacement and underlying dwell times 
%    \item In the absence of an external potential or complicated boundary conditions, 
%    one can define random walk based on independent hop length and dwell time 
%    distributions and arrive at any type of diffusive behavior. 
  \end{itemize}

  % BJC1: moved from methods
  \noindent Three well-known classes of behavior leading to anomalous subdiffusion are 
  continuous time random walks (CTRW), fractional Brownian motion
  (FBM) and random walks on fractals (RWF).\cite{meroz_toolbox_2015}
  \begin{itemize}
    \item The theory is frequently used to describe single particle trajectories
    \item A CTRW is characterized by a distribution of hop lengths and 
    dwell times, where each each hop is characterized by independent random draws from 
    each distribution.\cite{montroll_random_1965,morrin_three_2018}
    \item FBM is common in crowded, viscoelastic environments where each jump comes 
    from a Gaussian distribution but is anti-correlated to its previous 
    step.~\cite{mandelbrot_fractional_1968,jeon_fractional_2010,banks_anomalous_2005}
    \item An RWF is imposed by a system's geometry. Systems with tortuous pathways and dead
    ends cause anti-correlated motion.\cite{meroz_toolbox_2015,neusius_subdiffusion_2008}
    \item The processes described above can happen alone or in combination. 
    \item Meroz and Sokolov suggested a decision tree for determining which subdiffusive
    mechanisms apply to a system, adapted in Figure X. 
    \item Of course, more complicated variations on these models can arise which we 
    will incorporate as needed.	  % BJC1: e.g. fractional levy motion or langevin with levy noise
  \end{itemize}

@@

  While elegant, pure anomalous diffusion theory may in some cases give only a qualitative
  understanding of transport mechanisms.
  \begin{itemize}
    \item Complex temporal relationships in between hops may have a significant effect on the average MSD
    \item Inhomogeneity of a medium or multiple types of interactions may lead to a number of
    distinct diffusive regimes. 
    \item In this case, a state based model may be useful. 
  \end{itemize}

  \noindent We can use the observations of transport from our previous work in order to create 
  state-based models.
  \begin{itemize}
  	\item We know that dynamics differ between solutes moving inside the pores and 
  	in the tails.
  	\item We can use the anomalous diffusion approach to model to our data as a 
  	function of radial distance from the closest pore center.
  	\item Alternatively, we can abandon the anomalous diffusion theory and adapt the 
  	methods of Markov state modeling.
  \end{itemize}

  % moved parts of method to here.
  Markov state models (MSMs) are a popular class of models used to project long timescale
  behavior of molecules based on molecular simulation trajectories by identifying
  different dynamical modes and quantifying the rates of transitions between them.
  \begin{itemize}
    \item MSMs are frequently used to study systems with slow dynamics, such as protein 
    folding.~\cite{snow_how_2005,chodera_automatic_2007}
    \item Researchers typically aim to come up with a low dimensional 
    representation of the system based on features which preserve the process
    kinetics.
    \item This still often results in thousands of distinct states. 
  \end{itemize}

  In this work we attempt to model the long timescale behavior of the four 
  fastest moving solutes studied in our previous work.
  \begin{itemize}
    \item Specifically, we study methanol, urea, ethylene glycol and acetic acid
    \item In addition to moving quickly, these solutes have a range of chemical
    functionality and experience each of the three trapping mechanisms to different
    degrees.
    \item It is our goal to reproduce the MSDs exhibited by each solute on MD timescales
    and then to project the long-term dynamics of each solute. 
    \item We explore two modeling approaches
%  	\item The third uses an infinite hidden Markov model (iHMM), an unsupervised machine
%  	learning algorithm which groups solute behavior into different dynamical states.
  	% BJC: somewhere emphasize that this last approach is powerful since it doesn't 
  	% require anything different in systems with more complex geometries (i.e. QI)
  \end{itemize}

  Our first approach is based on the anomalous diffusion literature, applying the 
  mathematical formalism describing subdiffusive transport to observed solute 
  trajectories.
  \begin{itemize}
    \item We determine the most likely subdiffusive mechanisms at play using the 
    flow chart in figure X. 
    \item We then fit anomalous diffusion models in two ways: First, based on a
    single set of parameters. Second, we use a two-state approach incorporating
    two sets of parameters based on a solute's radial distance from the pore center. 
  \end{itemize}

  The second approach adapts the MSM framework to a relatively small set of known 
  states based on the three observed solute trapping mechanisms.
  \begin{itemize}
    \item We extend the framework to include dynamics between jumps. 
    \item That is we incorporate the magnitude and temporal dependence of the 
    time series fluctuations between each transition to a different state. 
    \item Temporal dependence has a significant influence on mean squared displacement.
  \end{itemize}

@@

%  We can use the observed mechanisms in order to choose a stochastic model of
%  sufficient complexity One can project long timescale solute behavior by modeling transport as a
%  stochastic process.
%  %MRS: as above, simple diffusion is stochastic, just not very complex.  So we want to add appropriate complexity.
%  \begin{itemize}
%    %MRS: best talk about these mechanisms in terms of the
%    %mathematical models they capture.  But I don't know that it's
%    %necessary to review all of them.  
%  
%    %It may be a better approach to start from the features of the
%    %problem (i.e. certain mechanisms) and choose the model (of the
%    %existing choices) that is appropriate for those mechanisms.
%    %I.e. diagnose first, then choose, rather than just applying a
%    %list of models useful in other field's timeseries. Let me know if that makes sense;
%    %I think how the methods applied are selected is a key to solve the problem.
%
%    \item Various disciplines study this type of problem in a number of ways.
%    \item In the simulation literature, most start by measuring mean squared 
%    displacement (MSD), as we did previously.  % This doesn't quite fit the thesis
%    \item Due to the short time scales most people can feasibly simulate, the MSD 
%    does not give a complete picture of what happens long-term, other than hinting
%    at types of mechanisms that could lead to the curve's shape.
%    %MRS: I'd say that MSD is too simple compared to what is going on mechanistically, but it's fundamentally the same idea.  
%    \item mean first passage time?
%    \item Markov state models are popular with researchers of proteins which have 
%    many dynamical states occurring over multiple time scales.
%    \item The anomalous diffusion literature explains the shape of MSD curves based
%    on rigorous mathematical representations of transport mechanisms. This field is
%    heavily used in analysis of single particle tracking experiments.
%    \item Finally, the financial literature tackles similar problems from a more 
%    general time series perspective. Their primary focus is towards predictions of
%    stock market returns and volatility but their methods are widely applicable to
%    modeling time series that exhibit a range of behavior.
%  \end{itemize}

@@

  We studied transport of solutes in the H\textsubscript{II} phase using an
  atomistic molecular model of four pores in a monoclinic unit cell with 
  10 \% water by weight. (See Figure tbd)
  \begin{itemize}
    \item Approximately one third of the water molecules occupy the tail region 
    with the rest near the pore center.
    \item We chose to study the 10 wt \% water system because solutes move 
    significantly faster than in the 5 wt \% system studied previously.
    \item Appropriate stochastic modeling requires that solutes sample the 
    accessible mechanisms with representative probability.  %BJC1: incorporated suggested phrasing
  \end{itemize}

@@

  % MRS: you intend this to be a separate model than the above?  If so, make clearer.
  % BJC1: renamed, is that what you meant?
  \subsection{The Anomalous Diffusion Model}\label{method:model_sFBM}

  Solutes in this system exhibit subdiffusive behavior, a type of anomalous diffusion.
  \begin{itemize}
  %MRS: you would probably want to be clear what the mechanistic reasons for for anomalous diffusion: 
  %want to talk about why, given the mechanisms you would consider. 
  %BJC1: could go through flow chart
  	\item During an anomalous diffusion process, the mean squared displacement (MSD)
  	does not grow linearly with time, rather it follows a power law of the form:
	\begin{equation} 
	\langle x^2(t) \rangle = K_{\alpha}t^\alpha
	\label{eqn:msd_form}
	\end{equation} 
	where $\alpha$ is the anomalous exponent and $K_\alpha$ is the
	generalized diffusion coefficient.
	\item A value of $\alpha < 1$ indicates a subdiffusive process, while values of
	$\alpha = 1$ and $\alpha > 0$ are characteristic of Brownian and superdiffusive
	motion respectively.
  \end{itemize}

@@

  % BJC: Should this be in the results? Lots of figures to show in support of the claims below
  % MRS: can you provide mechanistic reasoning for the below? Then can show the claims for the reasonable possibilities.
  % BJC1: modified this paragraph to parallel the flow chart. I'm still not sure the
  % best place for this. It seems like it belongs in the discussion, but I need to 
  % justify the choice in model so I can write the methods about the chosen models
  \noindent We believe that solutes in the system studied here exhibit subordinated 
  fractional Brownian motion (sFBM) where the parent process is FBM and the 
  leading process is a CTRW. 
  \begin{itemize}
    \item Working down the flow chart...
  	\item We observe non-stationary $z$-coordinate traces of each solute's
  	center of mass (COM). %BJC: figure of a z-coordinate trace in main text or in supporting info
  	\item Combined with the observation that the ensemble-averaged MSD differs
  	from the time-averaged MSD, we have evidence of non-ergodicity in this system, 
  	a trait inherent to CTRWs but not FBM or RWFs.~\cite{thiel_weak_2014}  % BJC: need a figure to show this
% BJC1: following two points are weak. Linearity of CTRW doesn't necessarily hold up on short time scales.
%  	\item For a pure CTRW, the time-averaged MSD should be linear.
%  	~\cite{neusius_subdiffusion_2008,meroz_subdiffusion_2010}
%  	\item However, a typical time-averaged solute MSD is sublinear (see supporting
%  	information), which suggests that there is another underlying subdiffusive mechanism.
	\item The process is likely subordinated to another process because the hop 
	lengths recorded after each dwell period are anti-correlated.
    % BJC1: need to eliminate RWF. I think a geometric argument should be convincing enough. Our
    % system doesn't have a fractal-like structure. Monomer positions fluctuate enough that
    % solutes can access any part of the membrane over time.
  	\item Given the viscoelastic nature of the monomers in our system, we believe
  	the hop lengths can be modeled with FBM. 
  	% BJC1 : probably won't need this equation. It's really just to validate my sfbm
  	% simulation procedure. Could go in supporting information with a plot showing that
  	% my sims fit to t^alpha*beta
 	\item For subordinated FBM, it can be shown that
  	\begin{equation}
  	\langle x^2(t) \rangle \simeq t^{\alpha\beta}
  	\end{equation}
  	where $\alpha$ is the anomalous exponent characteristic of the leading CTRW process
  	and $\beta$ is the anomalous exponent characteristic of the parent FBM process. 
  \end{itemize}

  \noindent We can characterize a CTRW process using the parameters which describe its
  dwell time and hop length distribution.  % This is different from the emissions of the markov state model where every single step is used to construct the distribution. 
  \begin{itemize}
	\item We used the \texttt{ruptures} python package in order to automatically identify
	changepoints in solute trajectories.\cite{truong_ruptures:_2018} (See Supporting
	Information for more details on chosen parameters. i.e. type of cost function, 
	cost function penalty tolerance, number of dimensions used)
	\item We used the corresponding hop lengths and dwell times between break points
	to construct empirical distributions.
  \end{itemize}

  % BJC1: not sure about this anymore. Might eventually change to Levy. But in the 
  % interest of simplicity, maybe leave it at Gaussian.
  For solutes in our system, the distribution of hop lengths appears to be
  well-represented by a Gaussian distribution.~\cite{metzler_random_2000,
  metzler_anomalous_2014,neusius_subdiffusion_2009}  % Note that this is different from the distribution of all increments (which is heavy-tailed)
  \begin{itemize}
	\item We are most interested in the standard deviation, $\sigma$, of the 
	hop length distribution.
  \end{itemize}

@@

  In general, we observe different dynamical behavior when solutes move inside the
  pore versus in the tail region.
  %MRS: could present the two models sequentiually.  Talk about one COULD have multiple modes, then in results present 1, then 2. 
  \begin{itemize}
    \item This inspired two models of varying complexity.
    \item We created a simple, single mode model with a single $\alpha$, $\sigma$ and $H$
    parameter fit to each solute.
    \item Our second, two mode model assigns a set of parameters to each of 2 modes based
    on the solute's radial location.
    \item Solutes in mode 1 are in the pore region defined as less than 0.75 nm from any
    pore center and all else are in mode 2, the tail region. 
    \item We determined this cut-off and described how we calculated radial 
    distance from the pore center in our previous work~\cite{coscia_chemically_2019}
  \end{itemize}

@@

  \subsection{Markov State Models}\label{method:MSMs}

  A Markov state model (MSM) decomposes a time series into a set of discrete states
  with transitions between states defined by a transition probability matrix, T,
  which describes the conditional probability of jumping to a specific 
  state given the previously observed state.~\cite{pande_everything_2010,wehmeyer_introduction_2018}
  \begin{itemize}
% BJC1: moved following to intro
%    \item In the context of molecular simulations, MSMs are frequently used
%    to study systems with slow dynamics, such as protein folding.~\cite{snow_how_2005,chodera_automatic_2007}
%    \item Researchers typically aim to come up with a low dimensional 
%    representation of the system based on features which preserve the process
%    kinetics. This facilitates the identification of discrete states from
%    which T is generated.
    \item Software packages such as MSMbuilder~\cite{beauchamp_msmbuilder2:_2011}
    and pyEMMA~\cite{scherer_pyemma_2015} provide work flows capable 
    of featurization and dimensional reduction.
  \end{itemize}

  In this work, we define a total of 8 discrete states based on the 3 trapping
  mechanisms observed in our previous work. 
  \begin{itemize}
  	\item Therefore, there is no need to subject our trajectories to any reduction
  	in dimensionality. 
  	\item However, we will adopt the validation techniques used in the packages
  	mentioned above.  
  	%BJC: not sure the best way to lay these out
  	\item The states we've chosen include all combinations of trapping mechanisms 
  	in the pore and out of the pore:
    \begin{table}[!htb]
	  \centering
	  \begin{tabular}{ll}
	  \hline
	  1. In tails                                  & 5. In pores                                  \\
	  2. In tails and hydrogen bonding             & 6. In pores and hydrogen bonding             \\
	  3. In tails and associated with sodium       & 7. In pores and associated with sodium       \\
	  4. In tails, hydrogen bonding and associated & 8. In pores, hydrogen bonding and associated \\
	  \hline
	  \end{tabular}
	  \caption{}\label{table:water_content}
   \end{table}
%  	\begin{enumerate}
%  	  \item Trapped in tails
%  	  \item Trapped in tails and hydrogen bonding
%  	  \item Trapped in tails and associated with sodium
%  	  \item Trapped in tails, hydrogen bonding and associated
%  	  \item In pores
%  	  \item In pores and hydrogen bonding
%  	  \item In pores and associated with sodium
%  	  \item In pores, hydrogen bonding and associated
%  	\end{enumerate}
  	\item These choices assume that there are no significant kinetic effects resulting
  	from solute conformational changes or pore size fluctuations, an assumption 
  	that may be relaxed in future work.
  \end{itemize}

  % BJC: does it make sense to use Dirichlet Process as a conjugate prior for
  % getting error on transition matrix?
  % BJC1: aha yes! \cite{bacallado_bayesian_2009}

  \noindent We constructed a state transition probability matrix based on observed solute
  trajectories.
  \begin{itemize}
    \item Using methods described in our previous work, we determined 	
    which, if any, trapping mechanisms affected the solute at each time step, and
    assigned the observation to a specific state according to the definitions 
    above.~\cite{coscia_chemically_2019}
    \item Based on the current and previous state observation, we incremented the
    appropriate entry of an $n~\times~n$ count matrix by 1, where $n$ is the number of states.
    \item For example, if we observed state 1 followed by a transition to state 3,
    we increment the $(1, 3)$ entry of the count matrix by 1.
    \item We generate the transition probability matrix from the count matrix by 
    normalizing the entries in each row so that they summed to unity.
  \end{itemize}
